\section*{Acknowledgments}

I will always remember the years of my graduate studies at the
Ben-Gurion University of the Negev as one of the most exciting
periods in my life. I take the  opportunity to thank the people who
led and accompanied me during this fascinating period.

Eyal Solomon Shimony, my teacher and advisor, introduced me to
the world of artificial intelligence research, and educated me in
research methodologies. He led me with patience through the years of
my studies, and his bright ideas and insightful remarks have always
been inspiring to me. Not just an excellent teacher, he has also
been an attentive listener, giving attention and showing care toward
every detail, and welcoming discussion of even half-baked ideas. I
only hope I could ever provide someone else with the same level of
guidance and support I received.

My research would not be possible without the education in computer
science which I received at the department. I am thankful to my
teachers who supplied me with systematic knowledge in a wide range of
areas, and would like to extend my thanks to Ronen Brafman, Amos
Beimel, Matya Katz, and Avraham Melkman, among others. As much as
from the courses, I learned from collaborative research, which I
was happy to conduct with Ronen Brafman, who also welcomed
informal presentation of my ideas on several occasions, and Ariel
Felner. 

The technical and administrative staff at the Computer Science
Department was always supportive and helpful. With their continuing
assistance I was able to concentrate on study, research, and
teaching.

Alexandra Polyan was the source of inspiration during an
extremely challenging period of my studies. Her ability to listen
to and discuss ideas from a wide range of areas is invaluable.

My research was partially supported by Israeli Science Foundation,
by the Lynne and William Frankel Center for Computer Sciences, and
by the Paul Ivanier Center for Robotics Research and Production
Management. 
